---
title: "IDS Final Presentation"
subtitle: "2022"
author: "Ben Mitchell, Eddie Jiang, Callum Knox, Sarah Brown-Scott"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(tidymodels)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# Load your data here
happiness <- read_csv("2015.csv")
countries <- read_csv("countries of the world.csv")
```

```{r include=FALSE}
colnames(happiness_countries)[4] <- "happiness_score"
colnames(happiness_countries)[6] <- "economy"
colnames(happiness_countries)[7] <- "family"
colnames(happiness_countries)[8] <- "health"
colnames(happiness_countries)[10] <- "trust"
colnames(happiness_countries)[9] <- "freedom"
colnames(happiness_countries)[20] <- "gdp"
```

```{r include=FALSE}
countries$Country[countries$Country == "Cote d\'Ivoire"] = "Ivory Coast"
countries$Country[countries$Country == "Bosnia & Herzegovina"] = "Bosnia and Herzegovina"
countries$Country[countries$Country == "Central African Rep."] = "Central African Republic"
countries$Country[countries$Country == "Congo, Dem. Rep."] = "Congo (Kinshasa)"
countries$Country[countries$Country == "Congo, Repub. of the"] = "Congo (Brazzaville)"
# error in kososvo
# and montenegro cont
```

```{r include=FALSE}
happiness_countries <- left_join(happiness, countries, by = "Country")
```


```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "img/confetti.jpg"
)
```

```{r include=FALSE}
set.seed(1111)

countries_split <- initial_split(happiness_countries, prop = 0.80)

train_data <- training(countries_split)
test_data  <- testing(countries_split)
```

```{r include=FALSE}
#model 1 with ROC curve and auc
countries_mod_1 <- linear_reg() %>%
  set_engine("lm")

countries_rec_1 <- recipe(happiness_score ~ economy + family + health + trust + freedom, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())

countries_wflow_1 <- workflow() %>%
  add_model(countries_mod_1) %>%
  add_recipe(countries_rec_1)

countries_fit_1 <- countries_wflow_1 %>%
  fit(data = train_data)

countries_pred_1 <- predict(countries_fit_1, test_data, type = "prob") %>%
  bind_cols(test_data)

countries_pred_1 %>%
  roc_auc(
    truth = happiness_score,
    .pred_Agree,
    event_level = "second"
  )

countries_pred_1 %>%
  roc_curve(
    truth = happiness_score,
    .pred_Agree,
    event_level = "second"
  ) %>%
  autoplot()
```

```{r include=FALSE}
#model 2 with roc curve and auc
countries_mod_2 <- linear_reg() %>%
  set_engine("lm")

countries_rec_2 <- recipe(happiness_score ~ gdp + family + Deathrate + trust + freedom, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())

countries_wflow_2 <- workflow() %>%
  add_model(countries_mod_1) %>%
  add_recipe(countries_rec_1)

countries_fit_2 <- countries_wflow_1 %>%
  fit(data = train_data)

countries_pred_2 <- predict(countries_fit_1, test_data, type = "prob") %>%
  bind_cols(test_data)

countries_pred_2 %>%
  roc_auc(
    truth = happiness_score,
    .pred_Agree,
    event_level = "second"
  )

countries_pred_2 %>%
  roc_curve(
    truth = happiness_score,
    .pred_Agree,
    event_level = "second"
  ) %>%
  autoplot()
```
class: center, middle

## How  Do a Range of Factors Affect Worldwide Happiness???

---

class: center, middle

# Dataset Information

The dataset we will be using for the project will be the World Happiness Report.
The dataset was found on the website Kaggle (https://www.kaggle.com/datasets/unsdsn/world-happiness)

The dataset uses the variables of Country, Region, Happiness rank, Happiness score, Standard error, Economy(GDP), Family, Health (Life Expectancy), Freedom, Trust (Government Corruption), Generosity and Dystopia Residual.

The dataset we are using was collected in 2015.

The dataset is used to help governments, organizations and civil society to inform their policy-making decisions.
The happiness rank and score data was collected by the Gallup World Poll.

The Gallup World Poll is said to be “the most comprehensive and farthest-reaching survey of the world”.

On the Gallop website it is stated that the survey connects with more than 99% of the world's adult population through annual, nationally representative surveys with comparable metrics across countries.


---

class: inverse, center, middle

# Reasons Behind Question Choice

---
class: center, middle

- We selected this question as mental health is a prominent issue in society and analysis of the data in this way could give some indicators of ways to improve the overall happiness of a country.


- The variables we think will best suit our question and hypotheses are, government trust, how the country is doing in terms of GDP and economy, the countries citizen freedom rating and the projected life expectancy of its citizens.
 And how these factors can vary independently and jointly with the countries happiness rating. 
If factors are proportional or inverse with each other. 


---

class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = family, y = happiness_score))+
  geom_jitter() +
  geom_smooth(method = "lm") +
  labs(title = "Happiness Score with Family",
       x = "Family",
       y = "Happiness Score")

```


---

class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = freedom, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm") +
  labs(title = "Happiness Score with Freedom",
       x = "Freedom",
       y = "Happiness Score")

```


---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = health, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm") +
  labs(title = "Happiness Score with Health",
       x = "Health",
       y = "Happiness Score")

```


---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = trust, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm") +
  labs(title = "Happiness Score with Government Trust",
       x = "Trust",
       y = "Happiness Score")

```


---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = economy, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm") +
  labs(title = "Happiness Score with Economy", 
       subtitle = "In terms of GDP",
       x = "Economy",
       y = "Happiness Score")

```

---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = economy, y = gdp))+
  geom_jitter()+
  geom_smooth(method = "loess") +
  labs(title = "Hapiness Score against GDP",
       x = "Economy Score",
       y = "GDP")

```

---

```{r echo=FALSE}
happiness %>%
  ggplot(data = happiness, 
         mapping = aes(x = Region))+
  geom_bar()
```

---

```{r echo=FALSE}
happiness_countries %>%
  ggplot(data = happiness_countries, 
         mapping = aes(x = happiness_score, y = Region.x))+
  geom_point()+
  labs(title = "Happiness Score shown by Region",
       x = "Happiness Score",
       y = "Region")
```

---

# Code

```{r boring-analysis}
# a boring analysis
ave <- mean(happiness_countries$happiness_score)
print(ave)
```

---

